Large and Infinitary Quotient Inductive-Inductive Types

### 2020-01-10 draft/paper.tex
paper

\newcommand{\ul}[1]{\underline{#1}}
\newcommand{\ulGamma}{\ul{\Gamma}}
\newcommand{\ulDelta}{\ul{\Delta}}
\newcommand{\ulgamma}{\ul{\gamma}}
\newcommand{\ulOmega}{\ul{\Omega}}
\newcommand{\uldelta}{\ul{\delta}}
\newcommand{\ulsigma}{\ul{\sigma}}
\newcommand{\ulepsilon}{\ul{\epsilon}}
\newcommand{\ult}{\ul{t}}
\newcommand{\ulu}{\ul{u}}
\newcommand{\ulA}{\ul{A}}
\newcommand{\ulB}{\ul{B}}
\newcommand{\tos}{\mathsf{tos}}

@@

\subsection{Universe}

@@

\begin{definition}[The $\Set$ model of $\ToS$] For each $j$, we have $\bS : \ToS_{j+1, j}$, which can be given by restricting the $\bM$ model
of Section \ref{sec:semantics} so that we only have the first $\Con$ components
in the interpretations for contexts, substitution, types, term, and we only have
actions on contexts in the interpretations of term and substitution
formers. Hence, we have:
\begin{alignat*}{3}
  & \Con_{\bS} && :\equiv \Set_{j+2} \\
  & \Sub_{\bS}\,\Gamma\,\Delta && :\equiv \Gamma \ra \Delta\\
  & \Ty_{\bS}\,\Gamma && :\equiv \Gamma \ra \Set_{j+2}\\
  & \Tm_{\bS}\,\Gamma\,A && :\equiv (\gamma : \Gamma) \ra A\,\gamma
\end{alignat*}
\end{definition}

For a syntactic signature $\Gamma$, $\llb\Gamma\rrb_{\bS}$ yields exactly the
type of $\Gamma$-algebras, e.g.\ $\llb\NatSig\rrb_{\bS}$ is a left-nested
$\Sigma$-type, which is the type of pointed sets with an endofunction.

\begin{definition}[Term algebra construction]
For each syntactic signature $\ulOmega : \SynSig_j$, we define a displayed $\ToS$
model over $\syn_j$, named $\bT$. The underlying sets are as follows:
\begin{alignat*}{3}
  & \Con\,\ulGamma :\equiv \Sub_{\syn}\,\ulOmega\,\ulGamma \ra \llb\ulGamma\rrb_{\bS}
\end{alignat*}

@@

%% write about limitations!

>>>>

\newcommand{\ul}[1]{\underline{#1}}
\newcommand{\ulGamma}{\ul{\Gamma}}
\newcommand{\ulDelta}{\ul{\Delta}}
\newcommand{\ulgamma}{\ul{\gamma}}
\newcommand{\ulOmega}{\ul{\Omega}}
\newcommand{\uldelta}{\ul{\delta}}
\newcommand{\ulsigma}{\ul{\sigma}}
\newcommand{\ulnu}{\ul{\nu}}
\newcommand{\ulepsilon}{\ul{\epsilon}}
\newcommand{\ult}{\ul{t}}
\newcommand{\ulu}{\ul{u}}
\newcommand{\ulA}{\ul{A}}
\newcommand{\ula}{\ul{a}}
\newcommand{\ulB}{\ul{B}}
\newcommand{\tos}{\mathsf{tos}}

@@

\subsection{Universe}
\label{sec:universe}

@@

\begin{definition}[The $\Set$ model of $\ToS$] For each $j$, we have $\bS : \ToS_{j+1, j}$, which can be given by restricting the $\bM$ model
of Section \ref{sec:semantics} so that we only have the first $\Con$ components
in the interpretations for contexts, substitution, types, term, and we only have
actions on contexts in the interpretations of term and substitution
formers. Hence, we have:
\begin{alignat*}{3}
  & \Con_{\bS} && :\equiv \Set_{j+2} \\
  & \Sub_{\bS}\,\Gamma\,\Delta && :\equiv \Gamma \ra \Delta\\
  & \Ty_{\bS}\,\Gamma && :\equiv \Gamma \ra \Set_{j+2}\\
  & \Tm_{\bS}\,\Gamma\,A && :\equiv (\gamma : \Gamma) \ra A\,\gamma
\end{alignat*}
Now, for $\Gamma : \Con_{\syn}$, the type of $\Gamma$-algebras is given by
$\llb\Gamma\rrb_{\bS}$. E.g.\, $\llb\NatSig\rrb_{\bS}$ yields a left-nested
$\Sigma$-type of pointed sets with an endofunction. Also, $\llb\Gamma\rrb_{\bM}$
extends $\llb\Gamma\rrb_{\bS}$ to an flCwF of $\Gamma$-algebras.
\end{definition}

\begin{definition}[Term algebra construction]
For each syntactic signature $\ulOmega : \SynSig_j$, we define a displayed $\ToS$
model over $\syn_j$, named $\bT$. The underlying sets are as follows:
\begin{alignat*}{3}
  & \Con_{\bT}\,\ulGamma &&:\equiv \Sub\,\ulOmega\,\ulGamma \ra \llb\ulGamma\rrb_{\bS}\\
  & \Sub_{\bT}\,\Gamma\,\Delta\,\ulsigma &&:\equiv (\ulnu : \Sub\,\ulOmega\,\ulGamma)
  \ra \Delta\,(\ulsigma\circ\nu) \simeq \llb\ulsigma\rrb_{\bS}\,(\Gamma\,\ulnu)\\
  & \Ty_{\bT}\,\Gamma\,\ulA &&:\equiv (\ulnu : \Sub\,\ulOmega\,\ulGamma) \ra
    \Tm\,\ulOmega\,(\ulA[\ulnu]) \ra \llb\ulA\rrb_{\bS}\,(\Gamma\,\ulnu)\\
  & \Tm_{\bT}\,\Gamma\,A\,\ult &&:\equiv (\ulnu : \Sub\,\ulOmega\,\ulGamma)\ra
    \llb\ulA\rrb_{\bS}\,\ulnu\,(\ult[\ulnu]) \simeq_{\id} \llb\ult\rrb_{\bS} (\Gamma\,\ulnu)
\end{alignat*}

Above, the $\simeq$ in the definition of $\Sub_{\bT}$ is a context isomorphism
in $\llb\ulDelta\rrb_{\bM}$, which is the flCwF of $\ulDelta$-algebras. The
$\simeq_{\id}$ in $\Ty_{\bT}$ is a vertical context isomorphism in the displayed
flCwF given by $\llb\ulA\rrb_{\bM}$.

So far, the underlying sets in $\bT$ are similar to what was given in
\cite{kaposi2019constructing} in the construction of term algebras, but there is
an important difference: in ibid.\, strict equalities are used instead of
isomorphisms. In our case, isomorphisms are necessary once again because of
infinitary functions types and our identity type; we shall see this shortly. The
universe is interpreted as follows:
\begin{alignat*}{3}
  & \U_{\bT} : (\ulnu : \Sub\,\ulOmega\,\ulGamma)(\ult : \Tm\,\ulOmega\,\U)\ra
              \Set_{j+1}\\
  & \U_{\bT}\,\ulnu\,\ult :\equiv \Tm\,\ulOmega\,(\El\,\ult)\\
  & \El_{\bT}\,a : (\ulnu : \Sub\,\ulOmega\,\ulGamma)
             (\ult : \Tm\,\ulOmega\,(\El\,(\ula[\ulnu])))\ra
              \llb\ula\rrb_{\bS}\,(\Gamma\,\ulnu)\\
  & \El_{\bT}\,a\,\ulnu\,\ult :\equiv (a\,\ulnu)\,\ult
\end{alignat*}

Hence, a syntactic $\ult : \Tm\,\ulOmega\,\U$ is interpreted as a set of terms
with type $\El\,\ult$. In the interpretation of $\El$, note that
\[
  a\,\ulnu : \U_{\bT}\,\ulnu\,(\ula[\ulnu]) \simeq_{\id}
                 \llb\ula\rrb_{\bS}\,(\Gamma\,\ulnu)
\]
hence
\[
a\,\ulnu : \Tm\,\ulOmega\,(\El\,(\ula[\ulnu])) \simeq_{\id}
           \llb\ula\rrb_{\bS}\,(\Gamma\,\ulnu)
\]
The $\simeq_{\id}$ above is just an isomorphism of sets, since it lives in
$\llb\U\rrb_{\bM}$ which was given as the flCwF of sets in Section
\ref{sec:universe}. This above isomorphism is a good summary of the
construction: the interpretation of a $\ula : \Tm\,\ulOmega\,\U$ in the term
algebra is isomorphic to a set of terms.

Inductive functions are interpreted by transport along such isomorphism:
\[
  \Pii_{\bT}\,a\,B\,\ulnu\,\ult :\equiv \lambda\,\alpha.\,
         B\,(\ulnu,\,(a\,\ulnu)^{-1}\,\alpha)\,(\ult\,\appitt\,((a\,\ulnu)^{-1}\,\alpha))
\]

For the infinitary function space, we need the following, where $\simeq_{\id}$ is
again set isomorphism.
\begin{alignat*}{3}
  & \Piinf_{\bT}\,A\,b\,\ulnu : \Tm\,\ulOmega\,(\El\,(\Piinf\,A\,(\lambda\,\alpha.\,(b\,\alpha)[\ulnu])))\\
  & \hspace{4.2em}\simeq_{\id} ((\alpha : A) \ra \llb\alpha\rrb_{\bS}\,(\Gamma\,\ulnu))
\end{alignat*}
This can be given using the natural isomorphism consisting of $\appinf$ and
$\laminf$. However, the sides are not strictly equal. For the identity type, we
build the following isomorphism using equality reflection.
\begin{alignat*}{3}
  & \Id_{\bT}\,a\,t\,u\,\ulnu : \Tm\,\ulOmega\,(\El\,(\Id\,(\ula[\ulnu])\,(\ult[\ulnu])\,(\ulu[\ulnu]))\\
  & \hspace{4.2em}\simeq_{\id} (\llb\ult\rrb_{\bS}\,(\Gamma\,\ulnu) =   \llb\ulu\rrb_{\bS}\,(\Gamma\,\ulnu))
\end{alignat*}

We omit the rest of the definition of $\bT$. The interpretations of equations in
the CwF and fl-structure parts become fairly technical, and we also need to
utilize iso-cleaving to interpret type substitution and substitution
laws. However, the basic shape of the model remains similar to
\cite{kaposi2019constructing}.

@@

%% write about limitations!
p
